---
title: Scatter-text Plots with Base R
author: Robert Schnitman
date: '2020-03-23'
slug: scatter-text-plots-with-base-r
categories:
  - R
tags:
  - plot
  - graphs
  - data analysis
  - Base R
---

To make scatter plots with text as points in Base R, we simply need to use `plot()`, set the scatter points to be white, and then plot the text with `text()`.

```{r}
# Trick R into not displaying points.
with(mtcars, plot(wt ~ mpg, pch = 1, col = 'white',
                  xlab = 'MPG',
                  ylab = 'Weight',
                  main = 'Weight vs. MPG'))

# Plot the labels on the graph.
with(mtcars, text(mpg, wt, row.names(mtcars), 
                  pos = 3, cex=0.0, col = 'cyan4'))
```

The function `plot_text()` below generalizes this process:

```{r}
plot_text <- function(y, x, labels, color = 'cyan4', ...) {
  
  # Blank out points
  plot(y ~ x, pch = 1, col = 'white', ...)
  
  # Plot the labels on the graph.
  text(x, y, labels, pos = 3, cex=0.0, col = color)
  
  
}
```

Now, let's apply it and add average lines with `abline()`.

```{r}
with(mtcars, plot_text(wt, 
                       mpg, 
                       row.names(mtcars), 
                       color = 'salmon',
                       ylab = 'wt',
                       xlab = 'mpg'))
with(mtcars, abline(h = mean(wt), v = mean(mpg), lty = 2))
```


And there you go!

---

# EDIT 2020-03-29
*You may also achieve similar results if you set `type = 'n'` in `plot()`, which retrospectively makes more sense than just setting the color to be white.*[^1]

[^1]: https://www.statmethods.net/graphs/line.html

```{r}
with(mtcars, plot(wt ~ mpg, pch = 1, type = 'n',
                  xlab = 'MPG',
                  ylab = 'Weight',
                  main = 'Weight vs. MPG'))

# Plot the labels on the graph.
with(mtcars, text(mpg, wt, row.names(mtcars), 
                  pos = 3, cex=0.0, col = 'cyan4'))
```