---
title: 'mop(): a New Way to sweep()'
author: Robert Schnitman
date: '2020-03-24'
slug: mop-a-new-way-to-sweep
categories:
  - R
tags:
  - afp
  - sweep
  - Data Analysis
  - Functional Programming
description: ''
topics: []
---

# Introduction

This blog post will compare `sweep()` and a function I've created called `mop()`. I argue that the latter is preferred over the former, as it is more concise in nature.

# The Old Way: `sweep`()

The function `sweep()`[^1] allows one to process data based on a summary statistic function--for example, dividing each element by a column's mean. A problem, however, arises: you are required to explicitly state the summary statistic value in the `STATS` input. So, taking our example of dividing each element by the respective column's average, we would have to do something along the lines of the following:

```{r, eval = FALSE}
sweep(mtcars, 2, sapply(mtcars, mean), `/`)
```

Because this way is quite verbose, I have written a function back in 2018 called `mop()`[^2] to handle this issue.

# The New Way: `mop()`

Essentially, `mop()` is a wrapper for `sweep(x, MARGIN, apply(...), FUN)`. This function is useful for indexing variables by their means, for example, so that the magnitude of a value relative to its average is known.

The four required arguments are `x`, `m`, `s`, and `f`--the collection (e.g. matrix), margin (1 for row-wise or 2 for column-wise), summary statistic function, and binary opertaor function, respectively. A fifth, optional argument `...` passes to `sweep()`. The output is typically a matrix or dataframe, depending on the inputs and functions being passed.

```{r}
mop <- function(x, m, s, f, ...) {
  
  # 1. Check inputs.
  f <- match.fun(f)
  s <- match.fun(s)
  
  diml <- length(dim(x))
  if (!diml) {

    stop('dim(x) must have a positive length. 
         Please make sure x is 2D!')
    
  }
  
  if (!any(m == 1:2)) {
   
    stop('The m (margin) input must either 
         be 1 (row-wise) or 2 (column-wise).') 
    
  }
  
  # 2. Sweep out the summary statistic function
  ## apply() allows us to control for margins.
  summ_stats <- apply(x, m, s) 
  
  output <- sweep(x, m, summ_stats, f, ...)
  
  # 3. Output should be 2D.
  ## its class should be the same as x.
  output
  
}
```

```{r mop1, cache = TRUE}
head(mop(mtcars, 2, mean, `/`))
# == head(sweep(mtcars, 2, apply(mtcars, 2, mean), `/`))
```

# Conclusion

In conclusion, `mop()` offers a superior version to `sweep()` by having the user pass a summary statistic *function* rather than an explicit summary statistic *value*. As such, the former function should be preferred over the latter one.

You can find this function and other functionals from my package, `afp` (Applied Functional Programming).[^3]

[^1]: https://www.r-bloggers.com/sweeping-through-data-in-r/
[^2]: https://github.com/robertschnitman/afp/blob/master/R/mop.R
[^3]: https://github.com/robertschnitman/afp