---
title: Mapping Techniques to Maintain Data Frame Consistency in Base R
author: Robert Schnitman
date: '2020-03-24'
slug: mapping-techniques-to-maintain-data-frame-consistency-in-base-r
categories:
  - R
tags:
  - Data Analysis
  - apply
  - map
  - data frames
  - Functional Programming
---

# Introduction
Maintaining data frame consistency within Base R can be difficult. The library `purrr`[^1] from the `tidyverse` solves this problem with its `map_df()` function. However, we can achieve similar results and expand upon them with base R functions. To do so, two methods will be used.

# Method 1: Use `lapply()`, `data.frame()`, and `do.call()`

To replicate `purrr`'s `map_df()`, we use three functions: `lapply()` to apply the function to some data; `data.frame()` to convert the output to a data frame; and `do.call()` to iteratively make said conversion.

[^1]: purrr.tidyverse.org/

```{r}
apply_df <- function(x, f, ...) {
  
  # 1. Apply the function. 
  ## Use other inputs if necessary.
  apply_it <- lapply(x, f, ...)
  
  # 2. Combine the elements into a data frame.
  output   <- do.call(data.frame, apply_it)
  
  # 3. Return the result.
  output
  
}
```

Let's test it out!

## The case of a single function

```{r}
apply_df(mtcars, mean)
```

## The case of two functions

What if we used two functions inside of `apply_df()`? With `purrr`'s `map_df`(), we would obtain two rows without rownames to identify the functions being used:

```{r}
# install.packages('purrr') # Install beforehand.

# Let's create a mean-SD function.
msd <- function(x) {
 
  c(mean = mean(x), sd = sd(x)) 
  
}

# Then let's use map_df() and msd() on mtcars.
purrr::map_df(mtcars, msd)
```

With our new `apply_df()` function, the rownames are maintained:

```{r}
apply_df(mtcars, msd)
```

# Method 2: Simply use `as.data.frame()` and `apply()`

What if we wanted to apply the function row-wise? Let's rewrite our function to take into account whether the user wants to apply a function either row-wise or column-wise.

```{r}
apply_df2 <- function(x, m, f, ...) {
  
  # 1. Apply the function. 
  ## Use other inputs if necessary.
  apply_it <- apply(x, m, f, ...) 
  
  # 2. Convert to data frame.
  ## This conversion depends on the "m" input.
  output   <- as.data.frame(apply_it)
  
  # 3. Return the result.
  output
  
}
```

Let's test it out! We'll output only the first three results for the following examples:

```{r}
apply_df2(mtcars, 1, msd)[, 1:3]
apply_df2(mtcars, 2, msd)[, 1:3]
```

# Conclusion

So, in conclusion, we can  maintain data-frame consistency in two ways: 

1. Combine `lapply()`, `data.frame()`, and `do.call()`; or

2. Simply use `as.data.frame()` over our `apply()` call.

The second method is useful for row-wise outputs. Try these functions out and expand upon them!