---
title: Using Residuals Percent in OLS Diagnostics
author: Robert Schnitman
date: '2020-03-23'
slug: using-residuals-percent-in-ols-diagnostics
categories:
  - R
tags:
  - Data Analysis
  - OLS
  - Ordinary Least Squares
  - regression
  - diagnostics
  - graphs
---

Many students (myself included) were taught to analyze the *raw* residuals when diagnosing regression models, but not in terms of *percent*. The benefit of the latter is that we can assess the relative magnitude of error from our regression model.

To display the residuals as a percent (henceforth *Residuals, %*), let's first load some necessary libraries.

```{r, message = FALSE, warning = FALSE}
libs <- c('tidyverse', 'magrittr', 'ggthemes', 'gridExtra')

# For each library, check if they are installed.
## If not, install and load them.
for (i in libs) {
  
  if (!require(i, character.only = TRUE)) {
    
    install.packages(i)
    library(i, character.only = TRUE)
    
  }
  
}
```

Second, let's estimate a model and generate the fitted values and residuals. We'll use the `mtcars` dataset pre-loaded into R for this example.

```{r}
# Estimate model.
## mtcars is pre-loaded into R.
mymodel <- lm(mpg ~ wt + hp + gear + am, mtcars)

# Generate fitted and residual values.
fr <- data.frame(fit = predict(mymodel),
                 res = resid(mymodel))

# Residuals as proportions.
fr$res_pct <- with(fr, res/(fit + res))
```

Third, let's generate a scatter plot between the residuals percents and the fitted values, as well as a histogram of the former.

```{r}
# Set up 1x2 plot.
par(mfrow = c(1, 2))

# Scatter plot
with(fr, plot(res_pct ~ fit,
              ylab = 'Residuals, %',
              xlab = 'Fitted Values',
              col  = 'cyan4',
              main = 'Residuals, % vs. Fitted Values'))
abline(h = 0, lty = 2)

# Histogram
hist(fr$res_pct, 
     col  = 'salmon',
     main = 'Distribution of Residuals %',
     xlab = 'Residuals, %')

```

We can replicate these graphs with `ggplot2` graphics.

```{r}
# Scatter plot
g1 <- ggplot(fr) + 
  aes(y = res_pct, x = fit) + 
  geom_point(col = 'cyan4') + 
  geom_hline(yintercept = 0, linetype = 2) + 
  labs(y = 'Residuals, %',
       x = 'Fitted Values',
       title = 'Residuals, % vs. Fitted Values') + 
  theme_calc() # From ggthemes.

# Histogram
g2 <- ggplot(fr) + 
  aes(x = res_pct) + 
  geom_histogram(col = 'black', 
                 fill = 'salmon',
                 bins = NROW(fr)/2) + 
  labs(x = 'Residuals, %',
       title = 'Distribution of Residuals %') + 
  theme_calc() # From ggthemes.

# Arrange the ggplots in a grid
grid.arrange(g1, g2, ncol = 2, nrow = 1) # From gridExtra.
```

In summary, we have learned how to calculate *Residuals, %* and graph them with Base R and `ggplot2`. Try using these techniques the next time you diagnose your models!