---
title: 'New Package and Book: dm, Statistical Data Management Tools for R'
author: Robert Schnitman
date: '2020-04-04'
slug: new-dm
categories:
  - R
  - Books
tags:
  - dm
  - data management
description: ''
topics: []
---

# Introduction

A couple years ago, [I started a writing a package on Github](https://github.com/robertschnitman/dm) that was inspired by the data managment functionalities in other statistical software such as Stata and SPSS. I got distracted by life, especially with work, and I practically stopped developing the package in 2019. This year, however, I finally sat down and finished developing this package, R documentation and all: the end-result was `dm`.

You can read the documentation on this package as a [Gitbook online](https://rs-dm.netlify.com/).

# Installation

Currently, this package is only available on Github, so please use `devtools` to install this package.

```{r, eval = FALSE}
if (!require(devools)) {
  
  install.packages('devtools')
  library(devtools)
  
}

install_github('robertschnitman/dm')
```

```{r, message = FALSE}
library(dm)
```


# Examples

## `recode()`

The `recode()` function presents a way to recode variables differently than [SPSS's method](https://libguides.library.kent.edu/SPSS/RecodeVariables).

```{r, warning = FALSE}
mtcars$am
recode(mtcars$am, 0:1, 2:3)
```

## `numNA()`

[As inspired by Stata](https://www.stata.com/manuals13/m-5missing.pdf), `numNA()` counts the number of missing values in a dataset; `rowNA()` counts the number of missing values by row; and `colNA()` counts the number of missing values by column.

[^1]: 

```{r}
numNA(airquality) # Total number of missing values.
rowNA(airquality) # Number of missing values by row.
colNA(airquality) # Number of missing values by column.
```

## `load_libraries`()

The `load_libraries()` function tests whether a set of libraries has been installed: if not, then it installs and loads them; if already installed, then the function loads the libraries as intended.

```{r, eval = FALSE}
load_libraries(c('tidyverse', 'ggformula', 'abind'))
```

# References

Kent State University. *SPSS Recode*. https://libguides.library.kent.edu/SPSS/RecodeVariables

Stata. *missing()*. https://www.stata.com/manuals13/m-5missing.pdf
